---
title: "Systematic review PRS code"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

## Packages

These packages are required for analysis: 

* tidyverse 
* tm
* SnowballC
* wordcloud
* RColorBrewer
* cowplot
* formattable
* choroplethr
* data.table
* viridis
* easyPubMed
* scales

```{r packages, include=FALSE}
library(tidyverse)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
library(cowplot)
library(formattable)
library(choroplethr)
library(data.table)
library(viridis)
library(easyPubMed)
library(scales)

```
## Import data

```{r import}
df <- read.csv(file = "C://Users//dkj201//Documents//Documents//A Canada//Dataextraction_basic.csv",
               header = TRUE,
               stringsAsFactors = FALSE)
colnames(df)[1] <- "ORN"

```

## Plots

# Number of papers by year

Number of papers containing a pharmacogenetic polygenic risk score by year of publication, as of our search in February 2021. Papers published in 2021 are not represented in the figure.  

```{r number_year}
years <- df$Year

years2 <- table(years)
str(years2)
years2 <- as.data.frame(years2)
years3 <- years2 %>%
  filter(years!="2021")

uomcol1 <- "#345D9D" #dark blue
uomcol2 <- "#00A0DF" #light blue

years_plot <- ggplot(years3, aes(x = years, y = Freq)) + 
  geom_bar(stat = "identity", colour = uomcol1, alpha = 0.8, fill = uomcol2,
           size = 1) +
  theme_bw() +
  labs(title = "Plot of number of papers by year (as of February 2021)",
       x = "",
       y = "Number of papers") +
  guides(fill = "none") +
  scale_y_continuous(limits = c(0,20))
years_plot
ggsave(filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//years.png",
       plot = years_plot,
       device = "png",
       width = 20,
       height = 15,
       units = "cm",
       dpi = 320)

```
# Disease area/ATC code/wordcloud

Number of papers included in the systematic review classified by ICD-10 Disease area. B) Number of papers classified by ATC classification of included drugs. ATC = Anatomical Therapeutic Chemical. ICD = International Classification of Diseases.

```{r demographics disease}
# A: Disease areas stacked barplot
disease <- data.frame(df$Disease.area, df$Year)
disease2 <- as_tibble(disease)
unique1 <- disease2$`df.Disease.area`

disease_year <- disease2 %>% 
  count(`df.Disease.area`, `df.Year`) %>% #adds a n count to each row
  spread(`df.Disease.area`, n, fill = 0) %>% #fills in blanks with zeroes
  mutate_at(vars(all_of(unique1)), cumsum) %>% #cumulative sum of each disease area
  gather(`df.Disease.area`, N, -`df.Year`, factor_key = T) %>% #gathers disease areas together with counts, ignores Year
  mutate(`df.Disease.area` = ordered(`df.Disease.area`, levels = rev(levels(`df.Disease.area`)))) #orders in reverse order
disease_year$df.Disease.area <- reorder(disease_year$df.Disease.area, -disease_year$N)
pal_disease <- brewer.pal(n = 8, "Set1")

disease_plot <- ggplot(disease_year, aes(`df.Year`, N, fill = `df.Disease.area`)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("Number of papers") +
  xlab("Year") +
  scale_fill_manual(values = pal_disease, name = "ICD-10 Disease Area") +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_y_continuous(limits = c(0,55), breaks = c(0,5,10,15,20,25,30,35,40,45,50,55)) +
  scale_x_continuous(breaks = c(2012,2014,2016,2018,2020))
disease_plot

ggsave(filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//diseaseplot.png", plot = disease_plot,
       device = "png",
       width = 20,
       height = 15,
       units = "cm",
       dpi = 320)
```

```{r demographics drugs}
# B: Drug areas stacked barplot

drugs2 <- data.frame(df$Intervention.area, df$Year)
unique1 <- drugs2$df.Intervention.area

drugs_year <- drugs2 %>% 
  count(`df.Intervention.area`, `df.Year`) %>% #adds a n count to each row
  spread(`df.Intervention.area`, n, fill = 0) %>% #fills in blanks with zeroes
  mutate_at(vars(all_of(unique1)), cumsum) %>% #cumulative sum of each disease area
  gather(`df.Intervention.area`, N, -`df.Year`, factor_key = T) %>% #gathers disease areas together with counts, ignores Year
  mutate(`df.Intervention.area` = ordered(`df.Intervention.area`, levels = rev(levels(`df.Intervention.area`)))) #orders in reverse order

#Change categories around
drugs_year$df.Intervention.area <- as.character(drugs_year$df.Intervention.area)
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "A Alimentary tract and metabolism"] <- "Alimentary tract and Metabolism (A)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "C Cardiovascular system"] <- "Cardiovascular system (C)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "N Nervous system (inc psychiatric)"] <- "Nervous system (inc psychiatric) (N)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "V Various"] <- "Various (V)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "H Systemic hormonal preparations (not sex or insulin)"] <- "Systemic hormonal preparations (H)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "L Antineoplastic and immunomodulating"] <- "Antineoplastic and immunomodulating (L)"
drugs_year$df.Intervention.area[drugs_year$df.Intervention.area == "J Antiinfectives for systemic use"] <- "Anti-infectives for systemic use (J)"

drugs_year$df.Intervention.area <- reorder(drugs_year$df.Intervention.area, -drugs_year$N)

pal_drugs <- brewer.pal(n = 8, "Set2")

drugs_plot <- ggplot(drugs_year, aes(`df.Year`, N, fill = `df.Intervention.area`)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("Number of papers") +
  xlab("Year") +
  scale_fill_manual(values = pal_drugs, name = "ATC code") +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_y_continuous(limits = c(0,55), breaks = c(0,5,10,15,20,25,30,35,40,45,50,55)) +
  scale_x_continuous(breaks = c(2012,2014,2016,2018,2020))
drugs_plot


ggsave(filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//drugsplot.png", plot = drugs_plot,
       device = "png",
       width = 20,
       height = 15,
       units = "cm",
       dpi = 320)
```

```{r demographics wordcloud}

# C: Wordcloud

words <- df$Drug.names
docs <- Corpus(VectorSource(words))

#Text transformation- removing special characters
toSpace <- content_transformer(function(x, pattern) gsub(pattern, " ", x))
docs<- tm_map(docs, toSpace, "/")
docs<- tm_map(docs, toSpace, "@")
docs<- tm_map(docs, toSpace, "\\|")

#Cleaning text 
excludeWords <- c("the", "containing", "specified", "others", "antipsychotics")
#go back and edit these once final word cloud is complete
docs <- tm_map(docs, removeNumbers)
docs <- tm_map(docs, removeWords, stopwords("english"))
docs <- tm_map(docs, removeWords, excludeWords)
docs <- tm_map(docs, removePunctuation)
docs <- tm_map(docs, stripWhitespace)

#Build term document matrix
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m), decreasing = TRUE)
d <- data.frame(word = names(v), freq=v)
head(d, 10)
topd <- head(d,10)


#word cloud
set.seed(123)
png(filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//drugs_wordcloud.png")
drugs_wordcloud <- wordcloud(words = d$word, freq = d$freq, min.freq = 2, 
          max.words = 100, random.order = TRUE, rot.per = 0.35,
          colors = brewer.pal(8, "Dark2"), scale = c(2,1)) #may need to run a few times to find a configuration that fits
dev.off()
```

Final figure is a combination of year_plot, disease_plot, and drugs_plot

```{r final fig3}

#cowplot the figures - including NULL to manage space between graphs
combo3 <- plot_grid(years_plot, NULL, disease_plot, NULL, drugs_plot, labels = c("A", "", "B", "","C"),
                    nrow = 1,
                    ncol = 5,
                    align = 'h',
                    rel_widths = c(1, 0.1, 1, 0.1, 1))
combo3
ggsave(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//combo3.png",
       plot = combo3,
       width = 90,
       height = 20,
       units = "cm",
       dpi = 320)
```
# Location of first and last authors


```{r pubmed queries}
my_query <- '33501934  33106254 32686723 32724131 33181205 33012170 32511779 32895570 32652002 32175767 32818716 32535905 32430334 30480086 31707832 31988971 32203155 32066683 30392411 31447353 30610379 30130674 30552317 29121268 30240446 28431800 29559929 29025760 28159590 26778674 26915630 26806281 24339138 23377640 31872049 32171635 31712719 31504448 30718812 33557049 30942838 30171181 33589590 24888364 31612640 31898914 30700811 31447354 28184875 26349998 29579042'
#not ideal but I found this works best, rather than importing the csv file
#Copy excel cols into word as plain text, find `enter` and replace with `space`
#copy result into my_query

# submit the query to PubMed to get PubMed IDs for articles
my_idlist <- get_pubmed_ids(my_query)

my_idlist$Count #check as expected

#Create an object which allows articles to be processed in batches of 70 articles
all_steps <- seq(1, my_idlist$Count, by = 70)

```

#first AUTHORS EXTRACTION FROM PUBMED

```{r function first authors}

#Extract data from pubmed, include only first authors 
all_author_results <- lapply(all_steps, function(i) {
  y <- fetch_pubmed_data(pubmed_id_list = my_idlist, retmax = 70, retstart = i)  
  yy <- table_articles_byAuth(y, included_authors = "first", getKeywords = TRUE)
  yy[, c("pmid", "year", "jabbrv", "lastname", "firstname", "address")]
})  
#this will take some time
```

```{r pubmed queries 2}
str(first_author_results) #some may be missing, will have to fill these in manually later

first_author_results2 <- do.call(rbind,first_author_results)
write.csv(first_author_results2, "C://Users//dkj201//Documents//Documents//A Canada//Final figures//first_author_results.csv", row.names=F)
#Manually create country details in Excel for speed, and add in details of missing papers
#copy to new sheet called first_author_results2.csv

#LAST AUTHORS EXTRACTION FROM PUBMED

#Extract data from pubmed, include only last authors 
last_author_results <- lapply(all_steps, function(i) {
  y <- fetch_pubmed_data(pubmed_id_list = my_idlist, retmax = 70, retstart = i)  
  yy <- table_articles_byAuth(y, included_authors = "last", getKeywords = TRUE)
  yy[, c("pmid", "year", "jabbrv", "lastname", "firstname", "address")]
})  
#this will take some time

str(last_author_results) #some may be missing, will have to fill these in manually later

last_author_results2 <- do.call(rbind,last_author_results)
write.csv(last_author_results2, "C://Users//dkj201//Documents//Documents//A Canada//Final figures//last_author_results.csv", row.names=F)
#Manually create country details in Excel for speed, and add in details of missing papers
#copy to new sheet called last_author_results2.csv

```

# Plotting locations of first and last authors

Location of A) first authors and B) last authors of the 51 papers included in our analysis. 

```{r plotting locations first authors}
#first AUTHORS

first_authors <- fread("C://Users//dkj201//Documents//Documents//A Canada//Final figures//first_author_results2.csv") #extracted from written sheet above

#Need to change the following countries
#bahrain, barbados, brunei darussalam, congo, faroe islands, hong kong, malta, 
#puerto rico, republic of macedonia, scotland, serbia, singapore, state of palestine, 
#syrian arab republic, n arab republic, tanzania
#Don't have any of these so can proceed without this step

#Countries not included
#andorra, bahrain, barbados, faroe islands, hong kong, puerto rico, singapore,state of palestine

#create counts
first_authors_count <- first_authors %>%
  count(pmid, lastname, allname)


#merge two files
first_authors_merge <- first_authors %>%
  left_join(first_authors_count, by = c("pmid", "lastname", "firstname"))


#add weight column to each affiliation - these are just 1 for ours
first_authors_merge$weight <- (1/first_authors_merge$n)


#create tables with weighted counts of countries
first_authors_countries <- aggregate(weight ~ country, first_authors_merge, sum)


#Rename all cols, make all countries lower case, value to numeric
colnames(first_authors_countries) <-c("region","value")
first_authors_countries <- data.frame(apply(first_authors_countries,2,tolower))
first_authors_countries$value <- as.numeric(first_authors_countries$value)
levels(first_authors_countries$region)[levels(first_authors_countries$region) == "united states"] <- "united states of america"
levels(first_authors_countries$region)[levels(first_authors_countries$region) == "the netherlands"] <- "netherlands"

#plot
first_authors_plot <- country_choropleth(first_authors_countries, title = "Location of first authors", 
                                         num_colors = 1) + 
  scale_fill_continuous(low="#eff3ff", high="#084594", na.value="seashell", limits = c(0,8),
                        name = "first Authors") +
  theme(plot.title = element_text(hjust = 0.5))
#warnings here are just for countries without data

first_authors_plot
ggsave(
  filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//first_authors_plot.png",
  plot = first_authors_plot,
  device = "png")


#crop off antarctica and white space manually
```

```{r plotting locations last authors}

#last AUTHORS

last_authors <- fread("C://Users//dkj201//Documents//Documents//A Canada//Final figures//last_author_results2.csv") #extracted from written sheet above

#Need to change the following countries
#bahrain, barbados, brunei darussalam, congo, faroe islands, hong kong, malta, 
#puerto rico, republic of macedonia, scotland, serbia, singapore, state of palestine, 
#syrian arab republicn arab republic, tanzania
#Don't have any of these so can proceed without this step

#Countries not included
#andorra, bahrain, barbados, faroe islands, hong kong, puerto rico, singapore,state of palestine

#create counts
last_authors_count <- last_authors %>%
  count(pmid, lastname, firstname)

#merge two files
last_authors_merge <- last_authors %>%
  left_join(last_authors_count, by = c("pmid", "lastname", "firstname"))

#add weight column to each affiliation - these are just 1 for ours
last_authors_merge$weight <- (1/last_authors_merge$n)

#create tables with weighted counts of countries
last_authors_countries <- aggregate(weight ~ country, last_authors_merge, sum)

#Rename all cols, make all countries lower case, value to numeric
colnames(last_authors_countries) <-c("region","value")
last_authors_countries <- data.frame(apply(last_authors_countries,2,tolower))
last_authors_countries$value <- as.numeric(last_authors_countries$value)
levels(last_authors_countries$region)[levels(last_authors_countries$region) == "united states"] <- "united states of america"
levels(last_authors_countries$region)[levels(last_authors_countries$region) == "the netherlands"] <- "netherlands"


#plot
last_authors_plot <- country_choropleth(last_authors_countries, title = "Location of last authors", 
                                         num_colors = 1) + 
  scale_fill_continuous(low="#eff3ff", high="#084594", na.value="seashell", limits = c(0,8),
                        name = "Authors") +
  theme(plot.title = element_text(hjust = 0.5))
#warnings here are just for countries without data

last_authors_plot
ggsave(
  filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//last_authors_plot.png",
  plot = last_authors_plot,
  device = "png")


#crop off antarctica and white space manually
```

# Plotting locations of all and last authors - combined

Location of all and last authors of the 51 papers included in our analysis. 

```{r plotting locations all authors}
#ALL AUTHORS

all_authors <- fread("C://Users//dkj201//Documents//Documents//A Canada//Final figures//all_author_results.csv") #extracted from written sheet above

#Need to change the following countries
#bahrain, barbados, brunei darussalam, congo, faroe islands, hong kong, malta, 
#puerto rico, republic of macedonia, scotland, serbia, singapore, state of palestine, 
#syrian arab republic, n arab republic, tanzania
#Don't have any of these so can proceed without this step

#Countries not included
#andorra, bahrain, barbados, faroe islands, hong kong, puerto rico, singapore,state of palestine

#create counts
all_authors_count <- all_authors %>%
  count(pmid, lastname, firstname)

#merge two files
all_authors_merge <- all_authors %>%
  left_join(all_authors_count, by = c("pmid", "lastname", "firstname"))


#add weight column to each affiliation
all_authors_merge$weight <- (1/all_authors_merge$n)
all_authors_countries <- all_authors_merge %>%
  select(country)

all_authors_countries2 <- data.frame(table(all_authors_countries))


#create tables with weighted counts of countries
all_authors_countries <- aggregate(weight ~ country, all_authors_merge, sum)


#Rename all cols, make all countries lower case, value to numeric
colnames(all_authors_countries2) <-c("region","value")
all_authors_countries2$region <- tolower(all_authors_countries2$region)
levels(all_authors_countries2$region)[levels(all_authors_countries2$region) == "united states"] <- "united states of america"
levels(all_authors_countries2$region)[levels(all_authors_countries2$region) == "the netherlands"] <- "netherlands"

#plot
all_authors_plot <- country_choropleth(all_authors_countries2, title = "Location of authors", 
                                         num_colors = 1) + 
  scale_fill_continuous(low="#eff3ff", high="#084594", na.value="seashell", 
                        name = "Authors",
                        limits = c(0,30)) + 
  theme(plot.title = element_text(hjust = 0.5)) 
  
#warnings here are just for countries without data

all_authors_plot
ggsave(
  filename = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//all_authors_plot.png",
  plot = all_authors_plot,
  device = "png")


#crop off antarctica and white space manually
```
# Tables of author locations

```{r location_tables_all}
#first authors
first_authors2 <- first_authors %>%
  select(country)


first_authors2 <- as.data.frame(table(first_authors2))

colnames(first_authors2) <- c("Country", "Frequency")

customBlue <- "#acb9ca"
customBlue0 <- "#b4c0cf"

first_authors2 <- arrange(first_authors2, desc(Frequency))
first_authors2$Country <- as.character(first_authors2$Country)
first_authors2[first_authors2 == "united states of america"] <- "USA"
first_authors2[first_authors2 == "netherlands"] <- "The Netherlands"
first_authors_table <- formattable(first_authors2,
            list(
              `Frequency` = color_bar(customBlue)
            ))
first_authors_table

#last authors
last_authors2 <- last_authors %>%
  select(country)

last_authors2 <- as.data.frame(table(last_authors2))

colnames(last_authors2) <- c("Country", "Frequency")

customBlue <- "#acb9ca"
customBlue0 <- "#b4c0cf"

last_authors2 <- arrange(last_authors2, desc(Frequency))
last_authors2$Country <- as.character(last_authors2$Country)
last_authors2[last_authors2 == "united states of america"] <- "USA"
last_authors2[last_authors2 == "netherlands"] <- "The Netherlands"
last_authors_table <- formattable(last_authors2,
            list(
              `Frequency` = color_bar(customBlue)
            ))
last_authors_table
                   
#ALL authors
all_authors2 <- all_authors %>%
  select(country)

all_authors2 <- as.data.frame(table(all_authors2))

colnames(all_authors2) <- c("Country", "Frequency")

customBlue <- "#acb9ca"
customBlue0 <- "#b4c0cf"

all_authors2 <- arrange(all_authors2, desc(Frequency))
all_authors2$Country <- as.character(all_authors2$Country)
all_authors2[all_authors2 == "united states"] <- "USA"
all_authors2[all_authors2 == "united kingdom"] <- "United Kingdom"
all_authors2[all_authors2 == "netherlands"] <- "The Netherlands"
all_authors_table <- formattable(all_authors2,
            list(
              `Frequency` = color_bar(customBlue)
            ))
all_authors_table                   
                   
```
# Ancestries of participants, by number of participants, BC

```{r ancestries.numbers}

ancest <- fread(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancestry_nums.txt",
                header = TRUE,
                colClasses = "numeric") #warnings are for columns that are not able to be coerced to numeric

bcyear <- ancest %>%
  select(Year, BC_AAC, BC_AF, BC_AME, BC_EAS, BC_AS, BC_EUR, BC_LAT, BC_NEA, BC_OCE, BC_SAS, BC_SSA, BC_NS)

#remove NAs
bcyear <- na.omit(bcyear)

#need to melt it tidy
molten_bcyear <- melt(bcyear, id.vars = "Year")

#BD code
bcyear <- molten_bcyear %>%
  rename(BC_ancestry = variable, BC_sample_size = value) %>%
  group_by(BC_ancestry) %>%
  filter(sum(BC_sample_size) > 0)
bcyear_all <- bcyear %>%
  group_by(Year, BC_ancestry) %>%
  summarise(BC_sample_size = sum(BC_sample_size)) %>%
  group_by(BC_ancestry) %>%
  mutate(cumulative_sum = cumsum(BC_sample_size))

#plotting
pal <- brewer.pal(n = 12, "Set3")

bcyear_plot <- ggplot(bcyear_all, aes(x = Year, y = cumulative_sum, fill = BC_ancestry)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("") +
  scale_fill_manual(values = pal, name = "BC Ancestry",
                    labels = c("African American\nAfro Caribbean",
                             "East Asian",
                             "European",
                             "Latino",
                             "Near Eastern",
                             "Central South Asian",
                             "Not specified")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        title = element_text(hjust = 0.5)) +
  labs(title = "Base cohort ancestries over time (numbers)")
bcyear_plot

ggsave("C://Users//dkj201//Documents//Documents//A Canada//Final figures//bcyear_plot_nums.png",
       plot = bcyear_plot)


```

Same plot, excluding European:
```{r ancestries.noeuropean.numbers}
ancest <- fread(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancestry_nums.txt",
                header = TRUE,
                colClasses = "numeric") #warnings are for columns that are not able to be coerced to numeric

bcyear.noeur <- ancest %>%
  select(Year, BC_AAC, BC_AF, BC_AME, BC_EAS, BC_AS, BC_LAT, BC_NEA, BC_OCE, BC_SAS, BC_SSA, BC_NS)

#remove NAs
bcyear.noeur <- na.omit(bcyear.noeur)

#need to melt it tidy
molten_bcyear.noeur <- melt(bcyear.noeur, id.vars = "Year")

#BD code
bcyear.noeur <- molten_bcyear.noeur %>%
  rename(BC_ancestry = variable, BC_sample_size = value) %>%
  group_by(BC_ancestry) %>%
  filter(sum(BC_sample_size) > 0)
bcyear_all.noeur <- bcyear.noeur %>%
  group_by(Year, BC_ancestry) %>%
  summarise(BC_sample_size = sum(BC_sample_size)) %>%
  group_by(BC_ancestry) %>%
  mutate(cumulative_sum = cumsum(BC_sample_size))

#plotting
pal <- brewer.pal(n = 12, "Set3")


bcyear_plot.noeur <- ggplot(bcyear_all.noeur, aes(x = Year, y = cumulative_sum, fill = BC_ancestry)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("") +
  scale_fill_manual(values = c("#9EDACF","#FFFFBE", "#FC9387", "#93BDDA","#FDBF79","#BEE37F"), 
                    name = "BC Ancestry",
                    labels = c("African American\nAfro Caribbean",
                             "East Asian",
                             "Latino",
                             "Near Eastern",
                             "Central South Asian",
                             "Not specified")
                    ) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        title = element_text(hjust = 0.5)) +
  labs(title = "Base cohort ancestries over time (numbers) without European populations")
bcyear_plot.noeur

ggsave(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//bcyear_plot_nums_noeur.png",
       plot = bcyear_plot.noeur)
```

# Ancestries of participants, by proportion of participants,BC

```{r ancestries.proportions.bc}
ancest <- fread(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancestry_nums.txt",
                header = TRUE,
                colClasses = "numeric") #warnings are for columns that are not able to be coerced to numeric

bcyear.prop <- ancest %>%
  select(Year, `BC sample size`, BC_AAC, BC_AF, BC_AME, BC_EAS, BC_AS, BC_EUR, BC_LAT, BC_NEA, BC_OCE, BC_SAS, BC_SSA, BC_NS) %>%
  rename(BC_sample_size = `BC sample size`)

bcyear.prop$BC_AAC <- bcyear.prop$BC_AAC/bcyear.prop$BC_sample_size
bcyear.prop$BC_AF <- bcyear.prop$BC_AF/bcyear.prop$BC_sample_size
bcyear.prop$BC_AME <- bcyear.prop$BC_AME/bcyear.prop$BC_sample_size
bcyear.prop$BC_EAS <- bcyear.prop$BC_EAS/bcyear.prop$BC_sample_size
bcyear.prop$BC_AS <- bcyear.prop$BC_AS/bcyear.prop$BC_sample_size
bcyear.prop$BC_EUR <- bcyear.prop$BC_EUR/bcyear.prop$BC_sample_size
bcyear.prop$BC_LAT <- bcyear.prop$BC_LAT/bcyear.prop$BC_sample_size
bcyear.prop$BC_NEA <- bcyear.prop$BC_NEA/bcyear.prop$BC_sample_size
bcyear.prop$BC_OCE <- bcyear.prop$BC_OCE/bcyear.prop$BC_sample_size
bcyear.prop$BC_SAS <- bcyear.prop$BC_SAS/bcyear.prop$BC_sample_size
bcyear.prop$BC_SSA <- bcyear.prop$BC_SSA/bcyear.prop$BC_sample_size
bcyear.prop$BC_NS <- bcyear.prop$BC_NS/bcyear.prop$BC_sample_size

#remove NAs
bcyear.prop <- na.omit(bcyear.prop)

#need to melt it tidy
molten_bcyear.prop <- melt(bcyear.prop, id.vars = "Year")

#BD code
bcyear.prop <- molten_bcyear.prop %>%
  rename(BC_ancestry = variable, BC_sample_size = value) %>%
  group_by(BC_ancestry) %>%
  filter(sum(BC_sample_size) > 0)
bcyear_all.prop <- bcyear.prop %>%
  group_by(Year, BC_ancestry) %>%
  summarise(BC_sample_size = sum(BC_sample_size)) %>%
  group_by(BC_ancestry) %>%
  mutate(cumulative_sum = cumsum(BC_sample_size)) %>%
  filter(BC_ancestry != "BC_sample_size") %>%
  arrange(cumulative_sum)

#plotting
pal <- brewer.pal(n = 12, "Set3")

#sort largest to smallest
bcyear_all.prop$BC_ancestry <- factor(bcyear_all.prop$BC_ancestry, levels = c(
  "BC_EUR", "BC_EAS", "BC_AAC", "BC_SAS", "BC_LAT", "BC_NEA", "BC_NS"
))

bcyear_plot.prop <- ggplot(bcyear_all.prop, aes(x = Year, y = cumulative_sum, fill = BC_ancestry)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("") +
  scale_fill_manual(values = pal, name = "BC Ancestry",
                    labels = c("European",
                               "East Asian",
                               "African American\nAfro Caribbean",
                               "Central South Asian",
                               "Latino",
                               "Near Eastern",
                               "Not specified")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        title = element_text(hjust = 0.5),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_y_continuous(limits = c(0,60)) +
  scale_x_continuous(breaks = c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020,2021)) +
  labs(title = "Base cohort ancestries over time (proportions)",
       y = "Proportion ancestry of total participants")
bcyear_plot.prop

ggsave("C://Users//dkj201//Documents//Documents//A Canada//Final figures//bcyear_plot_props.png",
       plot = bcyear_plot.prop)

```

Same plot, without European:
```{r ancestries.proportions.noeur}
ancest <- fread(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancestry_nums.txt",
                header = TRUE,
                colClasses = "numeric") #warnings are for columns that are not able to be coerced to numeric

bcyear.prop.noeur <- ancest %>%
  select(Year, `BC sample size`, BC_AAC, BC_AF, BC_AME, BC_EAS, BC_AS, BC_LAT, BC_NEA, BC_OCE, BC_SAS, BC_SSA, BC_NS) %>%
  rename(BC_sample_size = `BC sample size`)

bcyear.prop.noeur$BC_AAC <- bcyear.prop.noeur$BC_AAC/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_AF <- bcyear.prop.noeur$BC_AF/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_AME <- bcyear.prop.noeur$BC_AME/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_EAS <- bcyear.prop.noeur$BC_EAS/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_AS <- bcyear.prop.noeur$BC_AS/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_LAT <- bcyear.prop.noeur$BC_LAT/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_NEA <- bcyear.prop.noeur$BC_NEA/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_OCE <- bcyear.prop.noeur$BC_OCE/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_SAS <- bcyear.prop.noeur$BC_SAS/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_SSA <- bcyear.prop.noeur$BC_SSA/bcyear.prop.noeur$BC_sample_size
bcyear.prop.noeur$BC_NS <- bcyear.prop.noeur$BC_NS/bcyear.prop.noeur$BC_sample_size

#remove NAs
bcyear.prop.noeur <- na.omit(bcyear.prop.noeur)

#need to melt it tidy
molten_bcyear.prop.noeur <- melt(bcyear.prop.noeur, id.vars = "Year")

#BD code
bcyear.prop.noeur <- molten_bcyear.prop.noeur %>%
  rename(BC_ancestry = variable, BC_sample_size = value) %>%
  group_by(BC_ancestry) %>%
  filter(sum(BC_sample_size) > 0)
bcyear_all.prop.noeur <- bcyear.prop.noeur %>%
  group_by(Year, BC_ancestry) %>%
  summarise(BC_sample_size = sum(BC_sample_size)) %>%
  group_by(BC_ancestry) %>%
  mutate(cumulative_sum = cumsum(BC_sample_size)) %>%
  filter(BC_ancestry != "BC_sample_size")

#plotting
pal <- brewer.pal(n = 12, "Set3")

bcyear_plot.prop.noeur <- ggplot(bcyear_all.prop.noeur, aes(x = Year, y = cumulative_sum, fill = BC_ancestry)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  ylab("") +
  scale_fill_manual(values = c("#9EDACF","#FFFFBE", "#FC9387", "#93BDDA","#FDBF79","#BEE37F"),
                    labels = c("African American\nAfro Caribbean",
                             "East Asian",
                             "Latino",
                             "Near Eastern",
                             "Central South Asian",
                             "Not specified")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0,5)) +
  labs(title = "Base cohort ancestries over time (proportions) without European population")
bcyear_plot.prop.noeur

ggsave("C://Users//dkj201//Documents//Documents//A Canada//Final figures//bcyear_plot_props_noeur.png",
       plot = bcyear_plot.prop.noeur)
```

We will keep proportions of ancestries as final figure

#Proportion ancestries, TC

```{r ancestries.proportions.tc}
ancest <- fread(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancestry_nums.txt",
                header = TRUE,
                colClasses = "numeric") #warnings are for columns that are not able to be coerced to numeric

tcyear.prop <- ancest %>%
  select(Year, `TC sample size`, TC_AAC, TC_AF, TC_AME, TC_EAS, TC_AS, TC_EUR, TC_LAT, TC_NEA, TC_OCE, TC_SAS, TC_SSA, TC_NS) %>%
  rename(TC_sample_size = `TC sample size`)

tcyear.prop$TC_AAC <- tcyear.prop$TC_AAC/tcyear.prop$TC_sample_size
tcyear.prop$TC_AF <- tcyear.prop$TC_AF/tcyear.prop$TC_sample_size
tcyear.prop$TC_AME <- tcyear.prop$TC_AME/tcyear.prop$TC_sample_size
tcyear.prop$TC_EAS <- tcyear.prop$TC_EAS/tcyear.prop$TC_sample_size
tcyear.prop$TC_AS <- tcyear.prop$TC_AS/tcyear.prop$TC_sample_size
tcyear.prop$TC_EUR <- tcyear.prop$TC_EUR/tcyear.prop$TC_sample_size
tcyear.prop$TC_LAT <- tcyear.prop$TC_LAT/tcyear.prop$TC_sample_size
tcyear.prop$TC_NEA <- tcyear.prop$TC_NEA/tcyear.prop$TC_sample_size
tcyear.prop$TC_OCE <- tcyear.prop$TC_OCE/tcyear.prop$TC_sample_size
tcyear.prop$TC_SAS <- tcyear.prop$TC_SAS/tcyear.prop$TC_sample_size
tcyear.prop$TC_SSA <- tcyear.prop$TC_SSA/tcyear.prop$TC_sample_size
tcyear.prop$TC_NS <- tcyear.prop$TC_NS/tcyear.prop$TC_sample_size

#remove NAs
tcyear.prop <- na.omit(tcyear.prop)

#need to melt it tidy
molten_tcyear.prop <- melt(tcyear.prop, id.vars = "Year")

#BD code
tcyear.prop <- molten_tcyear.prop %>%
  rename(TC_ancestry = variable, TC_sample_size = value) %>%
  group_by(TC_ancestry) %>%
  filter(sum(TC_sample_size) > 0)
tcyear_all.prop <- tcyear.prop %>%
  group_by(Year, TC_ancestry) %>%
  summarise(TC_sample_size = sum(TC_sample_size)) %>%
  group_by(TC_ancestry) %>%
  mutate(cumulative_sum = cumsum(TC_sample_size)) %>%
  filter(TC_ancestry != "TC_sample_size") %>%
  arrange(cumulative_sum)


#plotting
pal <- brewer.pal(n = 12, "Set3")

#sort largest to smallest
tcyear_all.prop$TC_ancestry <- factor(tcyear_all.prop$TC_ancestry, levels = c(
  "TC_EUR", "TC_EAS", "TC_LAT", "TC_AS", "TC_AAC", "TC_AF", "TC_SAS", "TC_NS"
))

tcyear_plot.prop <- ggplot(tcyear_all.prop, aes(x = Year, y = cumulative_sum, fill = TC_ancestry)) +
  geom_area(position = "stack", alpha = 0.85) +
  theme_bw() +
  scale_fill_manual(values = c("#9EDACF",
                               "#FFFFBE",
                               "#93BDDA",
                               "forestgreen",
                               "#C8C4E0",
                               "red",
                               "#FC9387",
                               "#BEE37F"),
                    name = "TC Ancestry",
                    labels = c("European",
                               "East Asian",
                               "Latino",
                               "Asian",
                               "African American\nAfro Caribbean",
                               "African",
                               "Central South Asian",
                               "Not specified")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        title = element_text(hjust = 0.5),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_y_continuous(limits = c(0,50)) +
  scale_x_continuous(breaks = c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020,2021)) +
  labs(title = "Target cohort ancestries over time (proportions)",
       y = "Proportion ancestry of total participants")
tcyear_plot.prop

ggsave("C://Users//dkj201//Documents//Documents//A Canada//Final figures//tcyear_plot_props.png",
       plot = tcyear_plot.prop)

```

Into final figure using cowplot:

```{r cowplot ancestries}
ancest_combo <- plot_grid(bcyear_plot.prop, NULL, tcyear_plot.prop, labels = c("A", "", "B"),
                    nrow = 1,
                    ncol = 3,
          align = "h",
          rel_widths = c(1,0.1,1))
ancest_combo

ggsave(file = "C://Users//dkj201//Documents//Documents//A Canada//Final figures//ancest_combo.png",
       plot = ancest_combo,
       width = 45,
       height = 20,
       units = "cm",
       dpi = 320)

```

```{r sample sizes}
samplesize <- df %>%
  select(`BC.sample.size`, `TC.sample.size`, `Disease.area`)



```